---
title: "Vignette moveNT"
author: "Guillaume Bastille-Rousseau"
date: "April 18, 2017"
output:
  pdf_document: default
  html_document: default
---


``` {r eval=T, echo=F, message=F, warning=F}
library(devtools)
install_github("BastilleRousseau/moveNT")
library(moveNT)
library(adehabitatLT)
library(raster)
library(sp)
library(ade4)
library(adehabitatMA)
library(CircStats)
library(MASS)
library(boot)
library(moveHMM)
library(mclust)
library(igraph)
```

# Simulating movement strategies  - *sim_mov*

The function *sim_mov* generates movement trajectories including patches and movement between patches. Movement within patches can follow an Ornstein-Uhlenbeck process (based on *simm.mou* function from package *adehabitatLT*) or two-states movement model (based on *simmData* function from package *moveHMM*). Movement between patches is following a brownian bridge movement model (based on *simm.bb* function from package *adehabitatLT*). Generated outputs are of the class *ltraj* from package *adehabitatlt*.<br> <br>
\newline

``` {r eval=T}
# Simulating migration with two-states model 
mig<-sim_mov(type="2states", npatches=2, ratio=2, nswitch=25, ncore=150, grph=F)
mig
head(ld(mig))
plot(mig)

# Simulating multi-patches movement with Ornstein-Uhlenbeck process 
patches<-sim_mov(nswitch=25, ncore=150, ratio=5, type="OU", npatches=5, grph=T)

# Simulating sedentary movement
seden<-sim_mov(type="OU", npatches=10, spacecore=12, ratio=3, nswitch=150, ncore=20, grph=T)

```


# Converting movement to adjacency matrix  - *traj2adj*

The function *traj2adj* converts a trajectory object of class *ltraj* to an adjacency matrix. This is done by overlapping a grid over the relocation data and tallying the number of transitions among each pixel. Users need to specify the grid size, which can be based on distance travelled. The function *quant* is a wrapper that allows to sample a quantile of step length distribution from a *ltraj* object. Output produced by *traj2adj* is a list containing the adjacency matrix, the grid used (raster format), and a raster indicating pixel numbers that are occupied. These rasters are used by other functions such as *adj2stack* and *clustnet*. <br> <br>
\newline

``` {r eval=T}
# Using sedentary movement and user specific grid-size 
adj_seden<-traj2adj(seden, res=150) #Pixel size of 150m 
adj_seden[[1]] # Adjency matrix 
plot(adj_seden[[2]]) #Plot grid used

# Using multi-patches movement and median distance travelled 
adj_patches<-traj2adj(patches, res=quant(patches, p=0.5)) #Grid size based on median
dim(adj_patches[[1]]) # Size of the adjacency matrix 
plot(adj_patches[[2]]) #Plot grid used
plot(adj_patches[[3]]) #Plot occupied pixels 


# Using user defined grid
ras<-raster(nrows=10, ncols=10, xmn=0, ymn=0, xmx=6000, ymx=6000)
adj_patches2<-traj2adj(patches, res=quant(patches, p=0.5), grid=ras) #Grid size based on median
plot(adj_patches2[[2]]) #Crop version of the grid created 
```


# Calculation of network metrics  - *adj2stack*

The function *adj2stack* takes the output of function *traj2adj* and calculates a series of node- and graph-level metrics. Each metric is stored as a individual raster and the output is a raster stack combining each metric. Graph-level metrics are also stored as a raster, each containing an unique value. The function *graphmet* extracts graph-level metrics. The function *val* extracts only the occupied cells (remove NA) in a raster and allows the calculation of statistics from node-level metrics. <br> <br>
\newline  

``` {r eval=T}
# Using multi-patches movement and median distance travelled 
stck<-adj2stack(adj_patches,grph=T) #Plot the node-level metrics at the same time 
plot(stck) #Plot also the graph-level metrics (not really useful)
plot(stck[[3]]) #Plot only one metric (degree)
graphmet(stck) # Extract graph-level metrics 
cv(val(stck, 4)) #Extract coefficient of variation of node-level betweenness.  
```


# Clustering of node level metrics   - *clustnet*
The function *clustnet* applies a normal mixture model to node-level metrics in order to cluster them into separate groups (default = 2). The function takes the output of function *adj2stack* with the user specifying the metric to cluster and the number of groups. Return a list containing output of function *Mclust* from package *mclust* and a raster displaying classification. <br> <br>
\newline

``` {r eval=T}
# Using multi-patches movement and median distance travelled 
clust2<-clustnet(stck, id=3, nclust=2) # Clustering of degree in two groups
clust3<-clustnet(stck, id=4, nclust=3) #Clustering of betweenness in three groups 
summary(clust2[[1]])
plot(clust2[[2]])
summary(clust3[[1]])
plot(clust3[[2]])
```





